Inferring the Recombinant Genealogy of a Pandemic Pathogen in Real Time

### 2025-09-12 latex/main.tex
Updates from Overleaf

\subsubsection*{Comparison with UShER}
To facilitate comparison with the sc2ts ARG, we first converted the Viridian 
UShER tree from MAT~\cite{McBroome2021} to tskit format, which has similar
properties but is more general (see the section ``Ancestral Recombination
Graphs``).
We first downloaded the tree in JSON format (tree.all\_viridian.202409.jsonl)
from Figshare
and converted the ARG topology. We then downloaded the protobuf formatted
tree (tree.all\_viridian.202409.pb.gz), and exported the mutations to 
nucleotide format using \texttt{matUtils summary --translate}
(usher version 0.6.3), and translated
these mutations into tskit format. The resulting tszip file 
(usher\_viridian\_v1.0.trees.tsz) contains 5,345,019 nodes, 3,364,841
mutations and requires 29MB of storage space.
We validated that the data was correctly encoded by comparing the 
variant data for the UShER tskit tree sequence against our alignments.
There was an exact match between the non-ambiguous nucleotide calls 
at 27469 sites. At the remaining 39 sites there were a significant 
number of gap characters (deletion calls), suggesting that the mismatches
could be due to differences in the alignments used as input to UShER.
See notebook
\href{https://github.com/jeromekelleher/sc2ts-paper/blob/main/notebooks/validate\_usher\_tree.ipynb}{validate\_usher\_tree.ipynb}
for details.

We then computed intersection ARGs for sc2ts and UShER by 
``simplifying''\cite{Kelleher2018efficient,Wong2024} 
both to the 2,475,418 samples and 27,507 sites that are shared.

All conversions were performed as part of the reproducible 
postprocessing pipeline described in [section?]. The full UShER
tree and the intersection files are available for download on Zenodo
in tszip format.

@@

% See https://github.com/jeromekelleher/sc2ts-paper/issues/628
% Referenced by the main Results.
The ARG captures highly prevalent deletions in SARS-CoV-2 (Table~\ref{tab:major_dels}).

@@

Li et al.\cite{Li2023indels} performed a large-scale analysis of indels
using a global GISAID dataset (sampled up to August, 2022).
We examined the major indels reported in Table S1 of Li et al.\cite{Li2023indels}
which have a minimum frequency of 1\% (count of sequences divided by 9,149,680 sequences analyzed).
All of these indels are deletions and have high frequencies in the ARG
(sequence count divided by the total number of samples),
except a 9-base deletion at 11283, a 9-based deletion at 21987, and a 9-base deletion at 686.

>>>>

@@

% See https://github.com/jeromekelleher/sc2ts-paper/issues/628
% Referenced by the main Results.
The ARG captures highly prevalent deletions in SARS-CoV-2 (Table~\ref{tab:major_dels}).
Li et al.\cite{Li2023indels} performed a large-scale analysis of indels
using a global GISAID dataset (sampled up to August, 2022).
We examined the major indels reported in Table S1 of Li et al.\cite{Li2023indels}
which have a minimum frequency of 1\% (count of sequences divided by 9,149,680 sequences analyzed).
All of these indels are deletions and have high frequencies in the ARG
(sequence count divided by the total number of samples),
except a 9-base deletion at 11283, a 9-based deletion at 21987, and a 9-base deletion at 686.
During site remapping, the 9-base deletion at 11283 was split into
a 5-base deletion at 11283 (count of 339867) and a 4-base deletion at 11288 (count of 871782).

@@
